---
title: "Homework 10"
author: "Ryan Corydon and Emma Podlich"
date: "2023-04-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(plyr)
library(dplyr)
library(ggplot2)
library(maps)
```

## R Markdown

```{r}
accident <- "https://raw.githubusercontent.com/yumouqiu/DS202-Spring2023/main/Practice/data/fars2016/accident.csv"
accidentData = read.csv(accident)
View(accidentData)
```

## Data Background : Map Plot

```{r}
accidentData %>% ggplot(aes(x = LONGITUD, y = LATITUDE)) + geom_point(size = 0.05, colour = 'red') # Plot of All Data 

states <- map_data("state")
head(states)

states %>% ggplot(aes(x = long, y = lat)) + 
  geom_path(aes(group = group)) +
  geom_point(aes(x = LONGITUD, y = LATITUDE), size = 0.05, color = 'red', data = accidentData %>% filter(LONGITUD < 0, LONGITUD > -130)) # Plot with US Data

```

## Part 1: Accident Data

Are there some days of the week where more accidents happen than on others (use variable DAY_WEEK)?

```{r}

dailyAccident <- accidentData %>% group_by(DAY_WEEK) %>% summarise(count = n()) %>% ungroup()
dailyAccidentSorted <- dailyAccident %>% arrange(desc(count))
head(dailyAccidentSorted,7)

accidentData %>% group_by(DAY_WEEK) %>% tally()

```

```{r}
#The day of the week 7 has the most occurences of accidents of the week. Day 3 has the least.
```


What time of the day do accidents happen (use variable HOUR)?

```{r}
mean_value <- mean(accidentData$HOUR)
mean_value

#The average time an accident happens is at 13.4. If this is military hours you can assume most accidents are taking place around 1:00 PM.
```

What is the number of accidents with at least one drunk driver (use variable DRUNK_DR)?

```{r}
sum(accidentData$DRUNK_DR>=1)

#The number of accidents with at least 1 drunk driver is 8720 accidents within the data set.

```

## Part 2: Connecting Data

```{r}
person <- "https://raw.githubusercontent.com/yumouqiu/DS202-Spring2023/main/Practice/data/fars2016/person.csv"
personData <- read.csv(person)
View(personData)
```

Connect to the person table. Identify drivers (PER_TYP == 1, see fars manual ) and subset on them.

```{r}
subset_value <- personData[personData$PER_TYP==1, ]
View(subset_value)
```


Join accident and driver table (work out which variable(s) to use)

```{r}
newTable <- full_join(accidentData, subset_value)
```


Tally the number of accidents by day of the week (DAY_WEEK), hour of the day (HOUR) and gender (SEX). Visualize the results!

```{r}
Tally = newTable %>% group_by(DAY_WEEK, HOUR, SEX) %>% tally()
Tally

Tally %>% ggplot(aes(x = DAY_WEEK, y = n)) + geom_point(aes(colour = HOUR)) + facet_grid(SEX~ ., scale = "free_y") + geom_jitter() + ggtitle("Facet By SEX") + ylab("tally (n)")

```

